---
title: "Cumulative Biomass example"
author: "Matteo Zucchetta"
format: html
---

## Dependancies

loading required packages for running the example. For fitting the curves and estimating the parameters, only the `drc` package is needed.

```{r warning=F, message=F}
library(drc)
library(dplyr) # data handling
library(tidyr) # data handling
library(ggplot2) # plotting
library(ggthemes) # plotting
library(here) # handling paths
library(rfishbase) #for retrieving TLs from fishbase
```

## Loading the dataset

We use reconstructed catches from Sea Around Us (<https://www.seaaroundus.org/>) for the EEZs of Italy downloaded in March 24^th^ 2023.

```{r}
db <- read.csv(here('data/SAU EEZ 380,901,902 v50-0.csv'), sep = ",", dec = ".")


```

## Data handling

Normally you can use multiple local and specific sources to obtains trophic levels (TLs) for the species, but in this example we use the TLs from [fishbase](www.fishbase.org), retrieved with the `rfishbase` package.

First we extract the list of species in the dataset.

```{r message=F}
splist <- db |> group_by(scientific_name, common_name) |> summarise()
```

We use the function `ecology` from `rfishbase` to extract the TL

```{r message=F}
fb <- ecology(splist$scientific_name,
        fields=c("Species", "FoodTroph", "FoodSeTroph")) |> rename(scientific_name = Species)



```

Check for duplicates in retrived TLs

```{r}
fb %>% group_by(scientific_name) %>% filter(n() > 1)
```

We can remove the onw without TL information

```{r}
fb <- fb |> filter(!((scientific_name == "Sprattus sprattus") & is.na(FoodTroph)))
```

We obtain TLs for `r length(which(is.na(fb$FoodTroph)))` species out of `r nrow(splist)` taxa in the dataset. We will remove the taxa without TL for this examples, but this operation should be carefully evaluated in a real-world analysis.

We attach the TLs to the catches dataset and remove record without TL

```{r}
db <-  left_join(db, fb)
dat <- db |> filter(!is.na(FoodTroph ))
```

### Cumulating biomass

We group the TLs in TL classes, choosing a 0.1 step, and cumulating the biomass for each TL class, groupting for EEZ and year

```{r}
# TL step
step <- 0.1

#TL range
my.lim <- seq(1-step, 5.4 +step,by=step)

#TL labels (midpoint of the class)
my.lab <- my.lim + step/2


# factor 
dat$fTL <- cut(dat$FoodTroph , my.lim, labels=my.lab[-length(my.lim)])

# numeric label for the TL class
dat$fTLr <- as.numeric(levels(dat$fTL)[dat$fTL])

```

Cumulated biomass for TL class

```{r}

```
